---
title: "Predicting iris species using unsupervised and supervised machine learning"
author: "chenchi"
date: "1/25/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Summary
In this presentation, the `iris` data is analyzed using both unsupervised and supervised machine learning.
The following two algorithms are used to perform the whole analysis.

- **K-means clustering**
- **Decision tree**

A shiny app is developed to implement those two methods. The app features

- Model selection
- Feature selection
- Interactive plot of the k-means results
- Beautiful decision tree plot by `rpart.plot`

## K-means clustering

```{r, echo = TRUE, size= 'tiny'}
options(warn=-1)
suppressMessages(library(shiny))
suppressMessages(library(rpart))
suppressMessages(library(rpart.plot))
suppressMessages(library(ggplot2))
suppressMessages(library(plotly))
data <- iris[, c("Sepal.Width", "Sepal.Length")]
cluster_result <- kmeans(data, centers = 3)
iris['cluster'] <- as.factor(cluster_result$cluster)
prediction <- table(iris[, c('Species', 'cluster')])
clusters <- colnames(prediction)[max.col(prediction)]
clusters <- as.integer(clusters)
names(clusters) <- row.names(prediction)
iris['correct_predict'] <- clusters[iris[, 'Species']] == iris[, 'cluster']
accuracy = sum(iris[, 'correct_predict'])/dim(iris)[1]
g <- ggplot(iris, aes(x = Sepal.Width, y = Sepal.Length, color = Species, 
                          shape = 'correct_predict', size = 'correct_predict')) + geom_point() + 
                          ggtitle(paste('Accuracy ', accuracy))

```

## K-means (cont'd)

```{r, echo = TRUE}
ggplotly(g)
```

## Decision tree model

```{r, echo = TRUE, size = 'tiny', fig.align= 'center', fig.width=6, fig.height=3.5}
  fit <- rpart(Species ~., iris[, c('Species', "Sepal.Width", "Sepal.Length")])
  class.pred <- table(predict(fit, type="class"), iris$Species)
  accuracy <- sum(diag(class.pred))/sum(class.pred)
  rpart.plot(fit)
  title(paste('Accuracy ', accuracy))
```

